---
title: "Main script Schmits388143"
author: "Marie Schmit"
date: "2022-12-11"
output: html_document
---

```{r}
#install.packages("pheatmap")
#install.packages("caret")
```


Load libraries and packages
```{r}
require("mixOmics")
require("rgl")
require("matlab")
require("R.matlab")
require("ptw")
require("dplyr")
require("class")
require("gmodels")
require("caret")
require("rpart")
require("rpart.plot")
require("kernlab")
require("MASS")
require("grDevices")
require("factoextra")
require("cluster")
require("ggplot2")
require("RColorBrewer")
require("graphics")
require("caret")
require("mlr3")
library("pheatmap")
require("tidyverse")
require("mlbench")
require("mlr3viz")
require("mlr3learners")
require("randomForest")
require("mlr3extralearners")
require("devtools")
require("paradox")
require("mlr3tuning")
require("data.table")
require("randomForest")
require("reshape2")
require("Rmisc")
require("pls")
```
```{r}
#Install mlr3extralearners
devtools::install_local("C:/Programs/R-4.2.2/library/mlr3extralearners-0.5.49")
```



Functions are saved in a dedicated script.

```{r}
#Source the functions script
source("functions_MSchmit_s388143.R")
source("functions_regression_MSchmits388143.R", echo=TRUE)
```

# Data preparation and exploratory analysis
Load data files and create datasets.

```{r}
enose <- read.table("Enose_data-1.csv", sep = ",", header = TRUE, row.names = 1)
sensory <- read.table("Sensory_score-1.csv", sep = ",", header = TRUE, row.names = 1)
counts <- read.table("Bacterial_Counts-1.csv", sep = ",", header = TRUE, row.names = 1)
hplc <- read.table("HPLC_data-1.csv", sep = ",", header = TRUE, row.names = 1)
```


Combine all rows from the csv. Match the samples ID of HPLC and enose data with  sensory and bacterial counts data.
```{r}
# Merge enose and sensory
AllData_enose_sens <- merge_data(enose, sensory, AllData_enose_sens)

# Merge HPLC and sensory
AllData_HPLC_sens <- merge_data(hplc, sensory, AllData_HPLC_sens)

# Merge enose and counts
AllData_enose_counts <- merge_data(enose, counts, AllData_enose_counts)

#Merge HPLC and counts
AllData_HPLC_counts <- merge_data(hplc, counts, AllData_HPLC_counts)

#Merge sensory and counts
AllData_counts_sens <- merge_data(counts, sensory, AllData_counts_sens)

# Sensory values for enose and HPLC
AllData_enose_sens$sensory <- as.factor(AllData_enose_sens$sensory)
AllData_HPLC_sens$sensory <- as.factor(AllData_HPLC_sens$sensory)

sensory_enose <- AllData_enose_sens$sensory
sensory_HPLC <- AllData_HPLC_sens$sensory

# Samples ID
samples <- rownames(AllData_enose_counts)
```

## PCA visualisation
PCA of enose data grouped by sensors.
```{r}
pca.AllData_enose_sens <- pca_visualisation(pca.AllData_enose_sens, AllData_enose_sens, 3, sensory_enose, "lattice")

```
PCA of HPLC data grouped by sensors.
```{r}
pca.AllData_HPLC_sens <- pca_visualisation(pca.AllData_HPLC_sens, AllData_HPLC_sens, 3, sensory_HPLC, "lattice")

```
Create 3D PCA plots to better see all features.
```{r}
pca.AllData_enose_sens <- pca_visualisation(pca.AllData_enose_sens, AllData_enose_sens, 3, sensory_enose, "3d")
pca.AllData_HPLC_sens <- pca_visualisation(pca.AllData_HPLC_sens, AllData_HPLC_sens, 3, sensory_HPLC, "3d")
```

Create biplots of pca model.
```{r}
pca_var(pca.AllData_enose_sens, sensory_enose, AllData_enose_sens)
```

```{r}
pca_var(pca.AllData_HPLC_sens, sensory_HPLC, AllData_HPLC_sens)
```
HCA analysis
```{r}
hca_visualisation(AllData_enose_sens)
```
```{r}
hca_visualisation(AllData_HPLC_sens)
```

```{r}
pca_visualisation(pca.AllData_enose_sens, AllData_enose_sens, 3, sensory_enose, "lattice")
```

## 2. Bacterial counts against time for each temperature and bacteria

```{r}
# Add temperature and time columns to data frame counts of bacteria
# We remove the samples F1a and F1b because we do not know their temperature
counts_temp_time <- temp_time(AllData_counts_sens[-c(1,2),])
```
For each bacteria: TVC in full line, Pseudomonia in dot line.

```{r}
# Plot bacterial counts for each time for TVC
ggplot(data = counts_temp_time, 
       aes(x = as.numeric(counts_temp_time$Time), 
           y = as.numeric(counts_temp_time$TVC))) +
  #Color the lines by temperature
  geom_line(aes(colour = counts_temp_time$Temperature)) +
  #Add dashed lines for pseudomonads counts
  geom_line(aes(y = as.numeric(counts_temp_time$Pseudomonads),
                colour = counts_temp_time$Temperature), linetype = "dashed") +
  #Add labels
  xlab("Time (in hours)")+
  ylab("Bacterial count (Pseudomonads in dots)")+
  ggtitle("Bacterial count per time and temperature")+
  labs(color = "Temperature")

#Save plot
ggsave(filename = paste("Plots/2_CountsPerTime.png"))
```

## 3. Boxplots of bacterial counts for each type of bacteria against sensory score

```{r}
ggplot(data = counts_temp_time, mapping = aes())+
  #Boxes for TVC
  geom_boxplot(aes(group = counts_temp_time$sensory, 
                   y = counts_temp_time$TVC, x = counts_temp_time$sensory-0.25,
                   colour = "TVC"), width = 0.4)+
  #Boxes for pseudomonas
  geom_boxplot(aes(group = counts_temp_time$sensory, 
                   y = counts_temp_time$Pseudomonads, 
                   x = counts_temp_time$sensory+0.25, 
                   colour = "Pseudomonads"), width = 0.4)+
  #Create legend
  scale_color_manual("",
                     breaks = c("TVC", "Pseudomonads"),
                     values = c("Salmon", "Steelblue"))+
  # Set axis labels
  xlab("TVC counts")+
  ylab("Sensory")+
  ggtitle("Boxplots of bacterial counts against sensory")

#Save plot
ggsave(filename = paste("Plots/3_BoxPlot.png"))
```
# Classification
## Classification method 1: knn
### knn with HCLP
Partition the dataset with 7:3 optimisation
Partition the data once for HPLC.
```{r}
#Use the function Partition to create a partition of data
#Partition data from HPLC and sensory merged
part_HPLC_sens <- partition_data(AllData_HPLC_sens, 
                                 AllData_HPLC_sens$sensory, 0.7, 1)
```

## Knn classification
Perform classification for knn on HPLC data merged with sens.

```{r}
#Optimisation of k without scaling
optimal1 <- knn.optimisation(part_HPLC_sens$trainSet, part_HPLC_sens$testSet,
                             part_HPLC_sens$trainCl, part_HPLC_sens$testCl, 20, c())

optimal1

#Plot accuracy against time for non scaled data
plot(x = c(1:20), y = optimal1$k.accuracies, type='l', 
           main = "Accuracy by k without scale")

#Optimisation of scale methods and k values
optimal2 <- knn.optimisation(part_HPLC_sens$trainSet, part_HPLC_sens$testSet, part_HPLC_sens$trainCl, part_HPLC_sens$testCl, 20, c("center", "scale", "range"))

optimal2
```
Display the cross table for tunned model (k=7)
```{r}
cross_table_knn_HPLC <- cross_table_knn(part_HPLC_sens$trainSet,
                                        part_HPLC_sens$testSet,
                                        part_HPLC_sens$trainCl,
                                        part_HPLC_sens$testCl, k = 7)
#Display cross table
cross_table_knn_HPLC$cross.table$t
#Cross table, proportions
cross_table_knn_HPLC$cross.table$prop.tbl
```
Partition HPLC for 100 iterations, for tunned model knn (k=7).
```{r, include = FALSE}
#Run the operation knn
HPLC_sens_run_knn <- model.run(AllData_HPLC_sens, 
                               AllData_HPLC_sens$sensory, 
                               0.7, 100,
                               operation = function(trainSet, trainCl, testSet, testCl) {
            trainSet.knn <- trainSet[,1:(ncol(trainSet)-1)]
            testSet.knn <- testSet[,1:(ncol(testSet)-1)]
            #Apply knn model with k=7
            m <- knn(trainSet.knn, testSet.knn, trainCl, 7)
            return(m)
        })
```

```{r}
#Calculate cumulative mean
cum_mean_HPLC_knn <- cumulative.mean.accuracy(HPLC_sens_run_knn$accuracies)
#Display cumulative means for tuned knn model (k = 7), for 100 iterations
cum_mean_HPLC_knn
```
### Knn with enose
Perform classification for knn on HPLC data merged with sens.
Partition data
```{r}
# Partition data
part_enose_sens <- partition_data(AllData_enose_sens, AllData_enose_sens$sensory, .7, 1)
```

Model tunning
```{r}
#Optimisation of k without scaling
optimal1 <- knn.optimisation(part_enose_sens$trainSet, 
                             part_enose_sens$testSet, 
                             part_enose_sens$trainCl, 
                             part_enose_sens$testCl, 20, c())
#Accuracy of optimal model for knn, enose, without pre scaling
optimal1$bestAccuracy
#k value of optimal model
optimal1$bestk

#Plot accuracy against time for non scaled data
plot(x = c(1:20), y = optimal1$k.accuracies, type='l',
     main = "Accuracy by k without scale",
     xlab = "k",
     ylab = "Accuracy")

#Optimisation of scale methods and k values
optimal2 <- knn.optimisation(part_enose_sens$trainSet, 
                             part_enose_sens$testSet, 
                             part_enose_sens$trainCl, 
                             part_enose_sens$testCl, 
                             20, c("center", "scale", "range"))
#Accuracy of optimal model for knn, enose, with scaling
optimal2$bestAccuracy
#k of optimal model, enose, with prescaling
optimal2$bestk
#SCale of optimal model, enose, with prescaling
optimal2$bestScale
```
Partition enose for 100 iterations, for tunned model knn (k=3).
```{r, include = FALSE}
enose_sens_run_knn <- model.run(AllData_enose_sens, 
                                AllData_enose_sens$sensory, 
                                0.7, 100,
                                operation = function(trainSet, trainCl, testSet, testCl) {
            trainSet.knn <- trainSet[,1:(ncol(trainSet)-1)]
            testSet.knn <- testSet[,1:(ncol(testSet)-1)]
            #Apply knn model with k=7
            m <- knn(trainSet.knn, testSet.knn, trainCl, 7)
            return(m)
        })
```

Calculate cumulative means of previous model for 100 iterations
```{r}
#Calculate cumulative mean
cum_mean_enose_knn <- cumulative.mean.accuracy(enose_sens_run_knn$accuracies)
print(cum_mean_enose_knn)
```
## Classification method 2: svm - rd
### svm-rd with HCLP
```{r}
#List of kernels to test for calssification
kerList <- as.array(c('vanilladot', 'polydot', 'rbfdot', 'tanhdot'))

#Optimisation of the model: test of kernels from the list
optimal_svm_hplc <- svm.optimisation(part_HPLC_sens$trainSet, 
                                     part_HPLC_sens$testSet, 
                                     part_HPLC_sens$trainCl, 
                                     part_HPLC_sens$testCl, kerList)
#Accuracy, kernel, cross table and confusion matrix of tuned model
print("Accuracy:")
optimal_svm_hplc$bestAccuracy
print("Kernel:")
optimal_svm_hplc$bestKernel
print("Cross table:")
optimal_svm_hplc$bestCrossTable$t
print("Confusion matrix:")
optimal_svm_hplc$bestConfusionMatrix
```

```{r, include = FALSE}
#Run the optimised model for 100 iterations, save the resulting accuracies list
HPLC_sens_run_svm <- model.run(AllData_HPLC_sens, 
                               AllData_HPLC_sens$sensory, 
                               0.7, 100,
                               operation = function(trainSet, trainCl, testSet, testCl, kernel_type) {
            #Apply svm with kernel = rbfdot and 
            model_svm <- ksvm(sensory ~ ., data=trainSet, kernel = "rbfdot", C=1)
            kernel.predicted <- predict(model_svm, testSet, type="response")
            return(kernel.predicted)
        })
```


```{r}
#Calculate cumulative mean of previous model
cum_mean_HPLC_svm <- cumulative.mean.accuracy(HPLC_sens_run_svm$accuracies)
#Print cumulative mean accuracy of 100th iteration
cum_mean_HPLC_svm[100]
```


### svm-rd with enose
```{r}
#List of kernels to test for classification
kerList <- as.array(c('vanilladot', 'polydot', 'rbfdot', 'tanhdot'))

#Optimisation of the model: test of kernels from the list
optimal_svm_enose <- svm.optimisation(part_enose_sens$trainSet, part_enose_sens$testSet, part_enose_sens$trainCl, part_enose_sens$testCl, kerList)

#Print accuracy and kernel of tuned model
optimal_svm_enose$bestAccuracy
optimal_svm_enose$bestKernel
```

```{r, include = FALSE}
#Run the optimised model for 100 iterations, save the resulting accuracies list
enose_sens_run_svm <- model.run(AllData_enose_sens, 
                                AllData_enose_sens$sensory, 
                                0.7, 100,
                                operation = function(trainSet, trainCl, testSet, testCl, kernel_type) {
            #Apply svm with kernel = rbfdot and 
            model_svm <- ksvm(sensory ~ ., data=trainSet, kernel = "rbfdot", C=1)
            kernel.predicted <- predict(model_svm, testSet, type="response")
            return(kernel.predicted)
        })
```


```{r}
#Calculate cumulative mean
cum_mean_enose_svm <- cumulative.mean.accuracy(enose_sens_run_svm$accuracies)
#Print cumulative mean accuracy of 100th iteration
cum_mean_enose_svm[100]
```

## Classification method 3: Random Forest
### RF with data HPLC

Model building: Create model task
```{r}
#Create a learner
rf_class_learner_HPLC <- rf_class_learner(AllData_HPLC_sens)
rf_class_learner_HPLC$rf_task
#Check the number of elements for each sensory
rf_class_learner_HPLC$plt
```
Model tunning
```{r}
#List of hypoparameters
rf_class_learner_HPLC$learner$param_set$ids()
```

```{r}
#Pefrom tuning for hplc
rf_tuning_HPLC <- rf_tuning(ntree_min = 200, ntree_max = 500, 
                            mtry_min = 2, mtry_max = 19,
                            nodesize_min = 1, nodesize_max = 5, 
                            maxnodes_min = 2, maxnodes_max = 20, 
                            nb_evaluation = 20, part_ratio = .7, 
                            task_rf = rf_class_learner_HPLC$rf_task, 
                            AllData = AllData_HPLC_sens
                            )
```
Best hyperparameters for tuning rf learner, HPLC
```{r}
rf_tuning_HPLC$instance$result_learner_param_vals
```

Build final model with the tuned parameters for HPLC, rf
```{r}
learner_tun_HPLC <- built_model(learner_tun = rf_tuning_HPLC$learner_tun, 
                                instance = rf_tuning_HPLC$instance, 
                                task_train = rf_tuning_HPLC$task_train)
#Print tuned model
learner_tun_HPLC$model
```

Test final model
```{r}
test_HPLC_perf <- rf_test(rf_class_learner_HPLC$rf_task, 
                          learner_tun_HPLC, 
                          part_ratio = .7, 
                          split = rf_tuning_HPLC$split, 
                          title = "HPLC")
#Print accuracy
test_HPLC_perf$accuracy
#Print confusion matrix
test_HPLC_perf$conf_matr
#Plot performances
test_HPLC_perf$plt
```
Classification with rf, 100 iterations, HPLC
```{r}
#With 100 iterations, train and test the tuned model
rf_accuracies_HPLC <- run_rf(rf_class_learner_HPLC$rf_task, 
                             times = 100, 
                             learner = learner_tun_HPLC, 
                             ratio = .7)
#Calculate cumulative means accuracies
cum_mean_HPLC_rf <- cumulative.mean.accuracy(rf_accuracies_HPLC$accuracies)
#Cumulative mean accuracy of 100th iteration:
cum_mean_HPLC_rf[100]
```

### RF with enose
Model building: create model task
```{r}
#Create a learner for rf with enose data
rf_class_learner_enose <- rf_class_learner(AllData_enose_sens)
rf_class_learner_enose$rf_task
#Check the number of elements for each sensory
rf_class_learner_enose$plt
```
Model tunning, rf with enose data
```{r}
#List of hypoparameters
rf_class_learner_enose$learner$param_set$ids()
```

```{r}
#Perfrom tuning for rf, enose
rf_tuning_enose <- rf_tuning(ntree_min = 200, ntree_max = 500, 
                            mtry_min = 2, mtry_max = 8,
                            nodesize_min = 1, nodesize_max = 5, 
                            maxnodes_min = 2, maxnodes_max = 20, 
                            nb_evaluation = 20, part_ratio = .7, 
                            task_rf = rf_class_learner_enose$rf_task, 
                            AllData = AllData_enose_sens
                      )
```
Best hyperparameters
```{r}
#Display optimised hyperparameters
rf_tuning_enose$instance$result_learner_param_vals
```

Build final model with the tuned parameters
```{r}
learner_tun_enose <- built_model(learner_tun = rf_tuning_enose$learner_tun, 
                                 instance = rf_tuning_enose$instance, 
                                 task_train = rf_tuning_enose$task_train)
#Display tuned learner
learner_tun_enose$model
```

Test final model, tuned rf for enose data
```{r}
test_enose_perf <- rf_test(rf_class_learner_enose$rf_task, 
                           learner_tun_enose, 
                           part_ratio = .7, 
                           split = rf_tuning_enose$split, 
                           title = "enose")
#Print accuracy
test_enose_perf$accuracy
#Print confusion matrix
test_enose_perf$conf_matr
#Plot performances
test_enose_perf$plt
```
100 iterations, tuned rf for enose data
```{r}
#With 100 iterations, train and test the tunned model
rf_accuracies_enose <- run_rf(rf_class_learner_enose$rf_task, 
                              times = 100, 
                              learner = learner_tun_enose, 
                              ratio = .7)
#Calculate cumulative means accuracies
cum_mean_enose_rf <- cumulative.mean.accuracy(rf_accuracies_enose$accuracies)
#Cumulative mean accuracy for 100th iteration
cum_mean_enose_rf[100]
```
## 1. Bar plot of all cumulative mean accuracies, for 100 iterations

```{r}
#Create and save cumulative means plot for enose
cumulative_plot(cum_mean_enose_knn, cum_mean_enose_rf, cum_mean_enose_svm, "enose")
#Create and save cumulative means plot for HPLC
cumulative_plot(cum_mean_HPLC_knn, cum_mean_HPLC_rf, cum_mean_HPLC_svm, "HPLC")
```

## 2. Number and proportion of misclassified samples

```{r}
#Barplot of misclassifications proportion per class
list_methods <- list(HPLC_sens_run_knn$misclas$mis_prop_sum, 
                  HPLC_sens_run_svm$misclas$mis_prop_sum, 
                  rf_accuracies_HPLC$misclas$mis_prop_sum,
                  enose_sens_run_knn$misclas$mis_prop_sum, 
                  enose_sens_run_svm$misclas$mis_prop_sum, 
                  rf_accuracies_enose$misclas$mis_prop_sum
                  )
list_proportions <- c("HPLC_knn", "HPLC_svm", "HPLC_rf",
                      "enose_knn", "enose_svm", "enose_rf")

#Create and save barplots for every dataset and method
misclassified_proportion_barplot(list_methods, list_proportions)
```

##3. Importance of each variable for the prediction for one iteration
Variable importance for knn
```{r}
#Calculate the importance for knn
#HPLC data importances, kn
knn_var_importance_HPLC <- var_importance(AllData_HPLC_sens,
                                          AllData_HPLC_sens$sensory, .7, 1,
                                          sensory~., method = "knn", 
                                          tuneGrid = expand.grid(k = 1:20))
#Importance plot HPLC data, knn
importance_plot(knn_var_importance_HPLC$Imp, "HPLC, knn")

#Enose data importances, knn
knn_var_importance_enose <- var_importance(AllData_enose_sens,
                                           AllData_enose_sens$sensory, .7, 1,
                                           sensory~., method = "knn", 
                                           tuneGrid = expand.grid(k = 1:20))
#Importance plot enose data, knn
importance_plot(knn_var_importance_enose$Imp, "enose, knn")
```
Variable importance for svm
```{r}
#Calculate the importance for svm
#HPLC data importances, svm
svm_var_importance_HPLC <- var_importance(AllData_HPLC_sens,
                                          AllData_HPLC_sens$sensory, .7, 1,
                                          sensory~., method = "svmRadial", 
                                          tuneGrid = NULL)
#Importance plot, HPLC, svm
importance_plot(svm_var_importance_HPLC$Imp, "HPLC, svm")

#Enose data importances, svm
svm_var_importance_enose <- var_importance(AllData_enose_sens,
                                           AllData_enose_sens$sensory, .7, 1,
                                           sensory~., method = "svmRadial", 
                                           tuneGrid = NULL)
#Importance plot, enose, svm
importance_plot(svm_var_importance_enose$Imp, "enose, svm")
```
Overall variable importance for random forest
```{r}
#Return table of variables importance for HPLC
overall_rf_var_importance_HPLC <- var_importance(AllData_HPLC_sens,
                                          AllData_HPLC_sens$sensory, .7, 1,
                                          sensory~., method = "rf", 
                                          tuneGrid = expand.grid(mtry = 2:19))
#Importance plot
overall_importance_plot(overall_rf_var_importance_HPLC$Imp, "HPLC, rf")

#Variables importance for enose
overall_rf_var_importance_enose <- var_importance(AllData_enose_sens,
                                             AllData_enose_sens$sensory,0.7, 1,
                                             sensory~., method = "rf", 
                                             tuneGrid = expand.grid(mtry = 2:8))
#Importance plot
overall_importance_plot(overall_rf_var_importance_enose$Imp, "enose, rf")
```

Variable importance for random forest
```{r}
#Return table of variables importance for HPLC
rf_var_importance_HPLC <- rf_var_importance(AllData_HPLC_sens,
                                            AllData_HPLC_sens$sensory,
                                            0.7, 1,
                                            ntree = 200L, 
                                            mtry = 2L, 
                                            nodesize = 2L, 
                                            maxnodes = 20L)
#Importance plot
importance_plot(rf_var_importance_HPLC, "HPLC, rf")

#Variables importance for enose
rf_var_importance_enose <- rf_var_importance(AllData_enose_sens,
                                             AllData_enose_sens$sensory,0.7, 1,
                                             ntree = 200L, mtry = 2L, nodesize = 2L, maxnodes = 20L)
#Importance plot
importance_plot(rf_var_importance_enose, "enose, rf")
```

# Regression
## knn regression

Apply knn regression to HPLC data, to both bacteria types. Train the model to find the best parameters possible.

```{r}
#Apply tuning to Pseudomonads, HPLC
knn_tuning_HPLC_Pseud <- run_reg_model_tuning(AllData = AllData_HPLC_counts,
                                              perc_pred = .7, 
                                              times = 1, 
                                              method = "knn", 
                                              bacteria = "Pseudomonads",
                                              tuneGrid = expand.grid(k = 1:20), preProcess = c("center","scale"))

#Apply tuning to TVC, HPLC
knn_tuning_HPLC_TVC <- run_reg_model_tuning(AllData = AllData_HPLC_counts, 
                                            perc_pred = .7, 
                                            times = 1, 
                                            method = "knn", 
                                            bacteria = "TVC",
                                            tuneGrid = expand.grid(k = 1:20), preProcess = c("center","scale"))

#Display if preprocessing is necessary
knn_tuning_HPLC_Pseud$preProcess
knn_tuning_HPLC_TVC$preProcess
#Display best tunings
knn_tuning_HPLC_Pseud$model.fit$bestTune
knn_tuning_HPLC_TVC$model.fit$bestTune

```

Apply knn tuning to both bacteria types for enose data
```{r}
#Apply tuning to Pseudomonads, HPLC
knn_tuning_enose_Pseud <- run_reg_model_tuning(AllData = AllData_enose_counts,
                                               perc_pred = .7, 
                                               times = 1, 
                                               method = "knn", 
                                               bacteria = "Pseudomonads",
                                               tuneGrid = expand.grid(k = 1:20), preProcess = c("center","scale"))

#Apply tuning to TVC, HPLC
knn_tuning_enose_TVC <- run_reg_model_tuning(AllData = AllData_enose_counts,
                                             perc_pred = .7, 
                                             times = 1, 
                                             method = "knn", 
                                             bacteria = "TVC",
                                             tuneGrid = expand.grid(k = 1:20),
                                             preProcess = c("center","scale"))

knn_tuning_enose_Pseud$preProcess
knn_tuning_enose_TVC$preProcess
#Display best tunings
knn_tuning_enose_Pseud$model.fit$bestTune
knn_tuning_enose_TVC$model.fit$bestTune
```


Apply the tuned knn model to HPLC, for TVC and for 100 iteration
```{r, warning=FALSE}
knn_HPLC_TVC <- run_reg_model_iteration(AllData = AllData_HPLC_counts, 
                                        perc_pred = .7, 
                                        times = 100, 
                                        method = "knn", 
                                        bacteria = "TVC",
                                        tuneGrid = expand.grid(k = 1:20), 
                                        preProcess = c("center","scale"))
#Mean of RMSE for 100 iterations
knn_HPLC_TVC$mean_RMSE
knn_HPLC_TVC$sd_RMSE
knn_HPLC_TVC$CI_RMSE
```
Plot observed against predicted data for knn, HPLC, TVC
```{r}
plt <- plot_obs_pred(knn_HPLC_TVC$observed, 
                     knn_HPLC_TVC$predict.model, 
                     "Observed against predicted - KNN - HPLC TVC")

plt
```

knn model to HPLC, Pseudomonads, 100 iterations
```{r, warning=FALSE}
knn_reg_HPLC_Pseudo <- run_reg_model_iteration(AllData = AllData_HPLC_counts, 
                                               perc_pred = .7, 
                                               times = 100, 
                                               method = "knn", 
                                               bacteria = "Pseudomonads",
                                               tuneGrid = expand.grid(k = 1:20),
                                               preProcess = c("center","scale"))
#Mean of RMSE for 100 iterations
knn_reg_HPLC_Pseudo$mean_RMSE
#sd of RMSE
knn_reg_HPLC_Pseudo$sd_RMSE
#95% CI of RMSE
knn_reg_HPLC_Pseudo$CI_RMSE
```
Plot observed against predicted data for knn, HPLC, Pseudomonads
```{r}
plt <- plot_obs_pred(knn_reg_HPLC_Pseudo$observed, 
                     knn_reg_HPLC_Pseudo$predict.model, 
                     "Observed against predicted - KNN - HPLC pseudomonads")

plt
```

knn model to enose, TVC, 100 iterations
```{r, warning=FALSE}
knn_reg_enose_TVC <- run_reg_model_iteration(AllData = AllData_enose_counts, 
                                             perc_pred = .7, 
                                             times = 100, 
                                             method = "knn", 
                                             bacteria = "TVC",
                                             tuneGrid = expand.grid(k = 1:20), 
                                             preProcess = c("center","scale"))

#Mean of RMSE for 100 iterations
knn_reg_enose_TVC$mean_RMSE
#sd of RMSE
knn_reg_enose_TVC$sd_RMSE
#95% of RMSE
knn_reg_enose_TVC$CI_RMSE
```
Plot observed against predicted data for knn, enose, TVC
```{r}
plt <- plot_obs_pred(knn_reg_enose_TVC$observed, 
                     knn_reg_enose_TVC$predict.model, 
                     "Observed against predicted - KNN - enose TVC")

plt
```


knn model to enose, Pseudomonads, 100 iterations
```{r, warning=FALSE}
knn_reg_enose_Pseudo <- run_reg_model_iteration(AllData = AllData_enose_counts, 
                                                perc_pred = .7, 
                                                times = 100, 
                                                method = "knn", 
                                                bacteria = "Pseudomonads", 
                                                tuneGrid = expand.grid(k = 1:20),
                                                preProcess = c("center","scale"))
#Mean of RMSE for 100 iterations
knn_reg_enose_Pseudo$mean_RMSE
#sd of RMSE
knn_reg_enose_Pseudo$sd_RMSE
#95% CI of RMSE
knn_reg_enose_Pseudo$CI_RMSE
```
Plot observed against predicted data for knn, enose, pseudomonads
```{r}
plt <- plot_obs_pred(knn_reg_enose_Pseudo$observed, 
                   knn_reg_enose_Pseudo$predict.model, 
                  "Observed against predicted - KNN - enose pseudomonads")

plt
```

## Random forest regression

For HPLC, apply random forest (rf) regression, tuning for parameter mtry
#Apply tuning to Pseudomonads, HPLC
```{r}
rf_tuning_HPLC_Pseud <- run_reg_model_tuning(AllData = AllData_HPLC_counts, 
                                             perc_pred = .7, 
                                             times = 1, 
                                             method = "rf", 
                                             bacteria = "Pseudomonads",
                                             tuneGrid = expand.grid(mtry = 2:19),
                                             preProcess = c("center","scale"))

#Apply tuning to TVC, HPLC
rf_tuning_HPLC_TVC <- run_reg_model_tuning(AllData = AllData_HPLC_counts, 
                                           perc_pred = .7, 
                                           times = 1, 
                                           method = "rf", 
                                           bacteria = "TVC",
                                           tuneGrid = expand.grid(mtry = 2:19),
                                           preProcess = c("center","scale"))

#Display if preprocessing is necessary
rf_tuning_HPLC_Pseud$preProcess
rf_tuning_HPLC_TVC$preProcess
#Display best tunings
rf_tuning_HPLC_Pseud$model.fit$bestTune
rf_tuning_HPLC_TVC$model.fit$bestTune
```

Apply rf tuning to both bacteria types for enose data
```{r}
#Apply tuning to Pseudomonads, HPLC
rf_tuning_enose_Pseud <- run_reg_model_tuning(AllData = AllData_enose_counts, 
                                              perc_pred = .7, 
                                              times = 1, 
                                              method = "rf", 
                                              bacteria = "Pseudomonads",
                                              tuneGrid = expand.grid(mtry = 2:9),
                                              preProcess = c("center","scale"))

#Apply tuning to TVC, HPLC
rf_tuning_enose_TVC <- run_reg_model_tuning(AllData = AllData_enose_counts, 
                                            perc_pred = .7, 
                                            times = 1, 
                                            method = "rf", 
                                            bacteria = "TVC",
                                            tuneGrid = expand.grid(mtry = 2:9), 
                                            preProcess = c("center","scale"))

rf_tuning_enose_Pseud$preProcess
rf_tuning_enose_TVC$preProcess
#Display best tunings
rf_tuning_enose_Pseud$model.fit$bestTune
rf_tuning_enose_TVC$model.fit$bestTune
```

### 100 iterations

rf model to HPLC, for TVC and for 100 iteration
```{r, warning=FALSE}
run_rf_HPLC_TVC <- run_reg_model_iteration(AllData = AllData_HPLC_counts, 
                                           perc_pred = .7, 
                                           times = 100, 
                                           method = "knn", 
                                           bacteria = "TVC",
                                           tuneGrid = expand.grid(k = 2:19), 
                                           preProcess = c("center","scale"))
#Mean of RMSE for 100 iterations
run_rf_HPLC_TVC$mean_RMSE
#sd of RMSE
run_rf_HPLC_TVC$sd_RMSE
#CI of RMSE
run_rf_HPLC_TVC$CI_RMSE
```

Plot observed against predicted data for rf, HPLC, TVC
```{r}
plt <- plot_obs_pred(run_rf_HPLC_TVC$observed, 
                     run_rf_HPLC_TVC$predict.model, 
                     "Observed against predicted - RF - HPLC TVC")

plt
```

rf model to HPLC, pseudomonads, 100 iterations
```{r, warning=FALSE}
run_rf_HPLC_Pseudo <- run_reg_model_iteration(AllData = AllData_HPLC_counts, 
                                              perc_pred = .7, 
                                              times = 100, 
                                              method = "knn", 
                                              bacteria = "Pseudomonads",
                                              tuneGrid = expand.grid(k = 2:19), 
                                              preProcess = c("center","scale"))
#Mean of RMSE for 100 iterations
run_rf_HPLC_Pseudo$mean_RMSE
#sd of RMSE
run_rf_HPLC_Pseudo$sd_RMSE
#CI of RMSE
run_rf_HPLC_Pseudo$CI_RMSE
```

Plot observed against predicted data for rf, HPLC, pseudomonads
```{r}
plt <- plot_obs_pred(run_rf_HPLC_Pseudo$observed, 
                     run_rf_HPLC_Pseudo$predict.model, 
                     "Observed against predicted - RF - HPLC pseudomonads")

plt
```

rf model to enose, TVC, 100 iterations
```{r, warning=FALSE}
run_rf_enose_TVC <- run_reg_model_iteration(AllData = AllData_enose_counts, 
                                            perc_pred = .7, 
                                            times = 100, 
                                            method = "knn", 
                                            bacteria = "TVC",
                                            tuneGrid = expand.grid(k = 2:9), 
                                            preProcess = c("center","scale"))
#Mean of RMSE for 100 iterations
run_rf_enose_TVC$mean_RMSE
#sd of RMSE
run_rf_enose_TVC$sd_RMSE
#CI of RMSE
run_rf_enose_TVC$CI_RMSE
```

Plot observed against predicted data for rf, enose, TVC
```{r}
plt <- plot_obs_pred(run_rf_enose_TVC$observed, 
                     run_rf_enose_TVC$predict.model, 
                     "Observed against predicted - RF - enose TVC")

plt
```

rf model to enose, Pseudomonads, 100 iterations
```{r, warning=FALSE}
run_rf_enose_Pseudo <- run_reg_model_iteration(AllData = AllData_enose_counts, 
                                               perc_pred = .7, 
                                               times = 100, 
                                               method = "knn", 
                                               bacteria = "Pseudomonads",
                                               tuneGrid = expand.grid(k = 2:9), 
                                               preProcess = c("center","scale"))
#Mean of RMSE for 100 iterations
run_rf_enose_Pseudo$mean_RMSE
#sd of RMSE
run_rf_enose_Pseudo$sd_RMSE
#CI of RMSE
run_rf_enose_Pseudo$CI_RMSE
```

Plot observed against predicted data for rf, enose, pseudomonads
```{r}
plt <- plot_obs_pred(run_rf_enose_Pseudo$observed, 
                     run_rf_enose_Pseudo$predict.model, 
                     "Observed against predicted - RF - enose pseudomonads")

plt
```

## PLS regression

Apply partial least regression to HPLC data, to both bacteria types. Optimise, train and test the model.

### Tuning

```{r}
#Apply tuning to Pseudomonads, HPLC
plsr_tuning_HPLC_Pseud <- run_reg_model_tuning(AllData = AllData_HPLC_counts, perc_pred = .7, times = 1, method = "pls", bacteria = "Pseudomonads", 
                        tuneGrid = expand.grid(ncomp = 1:20), preProcess = c("center","scale"))

#Apply tuning to TVC, HPLC
plsr_tuning_HPLC_TVC <- run_reg_model_tuning(AllData = AllData_HPLC_counts, perc_pred = .7, times = 1, method = "pls", bacteria = "TVC", 
                        tuneGrid = expand.grid(ncomp = 1:20), preProcess = c("center","scale"))

#Display if preprocessing is necessary
plsr_tuning_HPLC_Pseud$preProcess
plsr_tuning_HPLC_TVC$preProcess
#Display best tunings
plsr_tuning_HPLC_Pseud$model.fit$bestTune
plsr_tuning_HPLC_TVC$model.fit$bestTune

```

Apply pls tuning to both bacteria types for enose data
```{r}
#Apply tuning to Pseudomonads, HPLC
pls_tuning_enose_Pseud <- run_reg_model_tuning(AllData = AllData_enose_counts, perc_pred = .7, times = 1, method = "pls", bacteria = "Pseudomonads", 
                        tuneGrid = expand.grid(ncomp = 1:10), preProcess = c("center","scale"))

#Apply tuning to TVC, HPLC
pls_tuning_enose_TVC <- run_reg_model_tuning(AllData = AllData_enose_counts, perc_pred = .7, times = 1, method = "pls", bacteria = "TVC", 
                        tuneGrid = expand.grid(ncomp = 1:10), preProcess = c("center","scale"))

pls_tuning_enose_Pseud$preProcess
pls_tuning_enose_TVC$preProcess
#Display best tunings
pls_tuning_enose_Pseud$model.fit$bestTune
pls_tuning_enose_TVC$model.fit$bestTune
```


Apply the tuned pls model to HPLC, for TVC and for 100 iteration
```{r, warning=FALSE}
pls_HPLC_TVC <- run_reg_model_iteration(AllData = AllData_HPLC_counts, 
                                        perc_pred = .7, 
                                        times = 100, 
                                        method = "pls", 
                                        bacteria = "TVC",
                                        tuneGrid = expand.grid(ncomp = 1:20), 
                                        preProcess = NULL)
#Mean of RMSE for 100 iterations
pls_HPLC_TVC$mean_RMSE
pls_HPLC_TVC$sd_RMSE
pls_HPLC_TVC$CI_RMSE
```
Plot observed against predicted data for knn, HPLC, TVC
```{r}
plt <- plot_obs_pred(pls_HPLC_TVC$observed, 
                     pls_HPLC_TVC$predict.model, 
                     "Observed against predicted - PLS - HPLC TVC")

plt
```

pls model to HPLC, Pseudomonads, 100 iterations
```{r, warning=FALSE}
pls_reg_HPLC_Pseudo <- run_reg_model_iteration(AllData = AllData_HPLC_counts, 
                                               perc_pred = .7, 
                                               times = 100, 
                                               method = "pls", 
                                               bacteria = "Pseudomonads",
                                               tuneGrid = expand.grid(ncomp = 1:20),
                                               preProcess = NULL)
#Mean of RMSE for 100 iterations
pls_reg_HPLC_Pseudo$mean_RMSE
#sd of RMSE
pls_reg_HPLC_Pseudo$sd_RMSE
#95% CI of RMSE
pls_reg_HPLC_Pseudo$CI_RMSE
```
Plot observed against predicted data for pls, HPLC, Pseudomonads
```{r}
plt <- plot_obs_pred(pls_reg_HPLC_Pseudo$observed, 
                     pls_reg_HPLC_Pseudo$predict.model, 
                     "Observed against predicted - PLS - HPLC pseudomonads")

plt
```

pls model to enose, TVC, 100 iterations
```{r, warning=FALSE}
pls_reg_enose_TVC <- run_reg_model_iteration(AllData = AllData_enose_counts, 
                                             perc_pred = .7, 
                                             times = 100, 
                                             method = "pls", 
                                             bacteria = "TVC",
                                             tuneGrid = expand.grid(ncomp = 1:10), 
                                             preProcess = NULL)

#Mean of RMSE for 100 iterations
pls_reg_enose_TVC$mean_RMSE
#sd of RMSE
pls_reg_enose_TVC$sd_RMSE
#95% of RMSE
pls_reg_enose_TVC$CI_RMSE
```
Plot observed against predicted data for knn, enose, TVC
```{r}
plt <- plot_obs_pred(pls_reg_enose_TVC$observed, 
                     pls_reg_enose_TVC$predict.model, 
                     "Observed against predicted - PLS - enose TVC")

plt
```


pls model to enose, Pseudomonads, 100 iterations
```{r, warning=FALSE}
pls_reg_enose_Pseudo <- run_reg_model_iteration(AllData = AllData_enose_counts, 
                                                perc_pred = .7, 
                                                times = 100, 
                                                method = "pls", 
                                                bacteria = "Pseudomonads", 
                                                tuneGrid = expand.grid(ncomp = 1:10),
                                                preProcess = NULL)
#Mean of RMSE for 100 iterations
pls_reg_enose_Pseudo$mean_RMSE
#sd of RMSE
pls_reg_enose_Pseudo$sd_RMSE
#95% CI of RMSE
pls_reg_enose_Pseudo$CI_RMSE
```
Plot observed against predicted data for pls, enose, pseudomonads
```{r}
plt <- plot_obs_pred(pls_reg_enose_Pseudo$observed, 
                   pls_reg_enose_Pseudo$predict.model, 
                  "Observed against predicted - PLS - enose pseudomonads")

plt
```